---
title: "Rolling Origin Cumulative and Monthly"
output: html_notebook
---

Expanding window:

```{r setup}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, comment= NA)

library(tidyverse)
library(tidyquant)
library(rsample)
library(yardstick)
library(broom)
library(timetk)

symbols <- c("SPY", "EFA", "IJS", "EEM", "AGG")


# The prices object will hold our daily price data.
prices <- 
  getSymbols(symbols, src = 'yahoo', 
             from = "2006-12-31",
             to = "2017-12-31",
             auto.assign = TRUE, 
             warnings = FALSE) %>% 
  map(~Ad(get(.))) %>% 
  reduce(merge) %>%
  `colnames<-`(symbols)


asset_returns_long <-  
  prices %>% 
  tk_tbl(preserve_index = TRUE, rename_index = "date") %>%
  gather(asset, returns, -date) %>% 
  group_by(asset) %>%  
  mutate(returns = (log(returns) - log(lag(returns)))) %>% 
  na.omit()

factors_data_address <- 
"http://mba.tuck.dartmouth.edu/pages/faculty/ken.french/ftp/Global_5_Factors_Daily_CSV.zip"

factors_csv_name <- "Global_5_Factors_Daily.csv"

temp <- tempfile()

download.file(
  # location of file to be downloaded
  factors_data_address,
  # where we want R to store that file
  temp, 
  quiet = TRUE)


Global_5_Factors <- 
  read_csv(unz(temp, factors_csv_name), skip = 6 ) %>%
  rename(date = X1, MKT = `Mkt-RF`) %>%
  mutate(date = ymd(parse_date_time(date, "%Y%m%d")))%>%
  mutate_if(is.numeric, funs(. / 100)) %>% 
  select(-RF)

data_joined_tidy <- 
  asset_returns_long %>%
  left_join(Global_5_Factors, by = "date") %>% 
  na.omit()


mod_form_1 <- as.formula(returns ~ MKT)
mod_form_2 <- as.formula(returns ~ MKT + SMB + HML)
mod_form_3 <- as.formula(returns ~ MKT + SMB + HML + RMW + CMA)

ff_rolling_flex <- function(split, formula) {
  
  split_for_data <- analysis(split)
  
  ff_model <- lm(formula, data = split_for_data)
  
  holdout <- assessment(split)
  
  rmse <-
    ff_model %>%
    augment(newdata = holdout) %>%
    rmse(returns, .fitted) %>%
    pull(.estimate)
  
}
```


JKR: let's go with A different asset altogether for the expanding and then for the monthly.

```{r}
efa_ff_data <- 
  data_joined_tidy %>% 
  filter(asset == "EFA")
```

```{r}
expanding_efa_splits <- rolling_origin(
  data       = efa_ff_data,
  initial    = 50,
  assess     = 19,
  cumulative = TRUE
)

expanding_efa_splits
```



```{r}
analysis(expanding_efa_splits$splits[[1]]) %>% 
  slice(c(1, n()))

# kept 2001-01-02, expanding training size
analysis(expanding_efa_splits$splits[[2]]) %>% 
  slice(c(1, n()))
```




```{r}
rmse_3_models_expanding <- 
  expanding_efa_splits %>% 
  mutate(model_1_rmse = map_dbl(expanding_efa_splits$splits,
                       ff_rolling_flex,
                       formula = mod_form_1),
         model_2_rmse = map_dbl(expanding_efa_splits$splits,
                       ff_rolling_flex,
                       formula = mod_form_2),
         model_3_rmse = map_dbl(expanding_efa_splits$splits,
                       ff_rolling_flex,
                       formula = mod_form_3))

rmse_3_models_expanding %>% 
  head()
```

```{r}
get_assess_start_date <- function(x) 
  min(assessment(x)$date)

get_analysis_start_date <- function(x) 
  min(analysis(x)$date)

rmse_3_models_expanding_with_start_date <-
  rmse_3_models_expanding %>%
  mutate(assess_start_date = map(expanding_efa_splits$splits, get_assess_start_date) %>% reduce(c),
         analysis_start_date = map(expanding_efa_splits$splits, get_analysis_start_date) %>% reduce(c)) %>%
  select(assess_start_date, analysis_start_date, everything())

rmse_3_models_expanding_with_start_date %>% 
  ggplot(aes(x = assess_start_date, y = model_1_rmse)) +
  geom_line()
```



Before we close, I know what you're thinking: hey, what if I don't want to split my data by number of days! I want to split my data by months and I can't use 30 days! Not to fear, `rsample` + `tibbletime` has us covered there too. We first create a new column to hold the month of each observation using `collapse_index(date, "monthly")`, then `group_by()` that column and `nest()` our data.  

```{r}
library(tibbletime)

  data_joined_tidy %>% 
  filter(asset == "SPY") %>% 
  mutate(nest_date = tibbletime::collapse_index(date, "monthly")) %>%
  group_by(nest_date) %>%
  nest() 
```

 Note that we didn't lose our original daily data, we just nested it. We could reaccees it by `unnest()`ing our data: 
 
```{r}
data_joined_tidy %>% 
  filter(asset == "SPY") %>% 
  mutate(nest_date = tibbletime::collapse_index(date, "monthly")) %>%
  group_by(nest_date) %>%
  nest(-nest_date)  %>% 
  unnest(data)
```
 
Now we want to create splits based on those months. Let's use the previous 3 months to train our model and then assess its performance on the next 1 month.

```{r}
spy_rolled_split_by_month <- 
  data_joined_tidy %>% 
  filter(asset == "SPY") %>% 
  mutate(nest_date = tibbletime::collapse_index(date, "monthly")) %>%
  group_by(nest_date) %>%
  nest() %>% 
  rolling_origin(initial = 3,
                 assess =  1,
                 cumulative = FALSE)

spy_rolled_split_by_month_cum <- 
  eem_for_roll %>% 
  mutate(nest_date = tibbletime::collapse_index(date, "monthly")) %>%
  group_by(nest_date) %>%
  nest() %>% 
  rolling_origin(initial = 3,
                 assess =  1,
                 cumulative = TRUE)


spy_rolled_split_by_month
```

Now we could pipe that `sply_rolled_split_by_month` object into a similar  model flow from above. 


```{r}
ff_three_rmse_monthly <- function(split){
  
analysis_set <- 
analysis(split)$data %>% 
bind_rows()

ff_model <- lm(returns ~ MKT + SMB + HML, data = analysis_set)
  
holdout <- 
  assessment(split)$data %>% 
  as.data.frame()
  
rmse <-
    ff_model %>%
    augment(newdata = holdout) %>%
    rmse(returns, .fitted) %>%
    pull(.estimate)
}

ff_three_rmse_monthly(spy_rolled_split_by_month$splits[[1]]) %>% 
  print()
```


```{r}
rmse_one_model_monthly_cum <- 
  # can handle cumulative too
  spy_rolled_split_by_month_cum %>% 
  mutate(model = map_dbl(
    .x = splits,
    .f = ~(ff_three_rmse_monthly(.x))))

rmse_one_model_monthly_cum %>% 
  head()
```


```{r}
ff_rolling_monthly_flex <- function(split, formula) {
  
analysis_set <- 
analysis(split)$data %>% 
bind_rows()
  
ff_model <- lm(formula, data = analysis_set)
  
holdout <- 
  assessment(split)$data %>% 
  as.data.frame()
  
  rmse <-
    ff_model %>%
    augment(newdata = holdout) %>%
    rmse(returns, .fitted) %>%
    pull(.estimate)
  
}

```


```{r}
rmse_3_model_monthly <- 
  spy_rolled_split_by_month %>% 
  mutate(model_1_rmse = map_dbl(spy_rolled_split_by_month$splits,
                       ff_rolling_monthly_flex,
                       formula = mod_form_1),
         model_2_rmse = map_dbl(spy_rolled_split_by_month$splits,
                       ff_rolling_monthly_flex,
                       formula = mod_form_2),
         model_3_rmse = map_dbl(spy_rolled_split_by_month$splits,
                       ff_rolling_monthly_flex,
                       formula = mod_form_3))

rmse_3_model_monthly %>% 
  head()
```

