---
title: "R Notebook"
output: html_notebook
---


```{r setup}
library(tidyquant)
library(tibbletime)
library(mixtools)
```


```{r}
spy_returns <- 
  tq_get("SPY",
         from = "1990-01-01") %>% 
  select(date, adjusted) %>% 
  mutate(returns = log(adjusted) - log(lag(adjusted))) %>% 
  select(-adjusted) %>% 
  na.omit()
```


```{r}
library(tidyquant)
library(tibbletime)
library(mixtools)
library(sigmoid)

mix_mod <- normalmixEM(spy_returns$returns, k = 2, lambda = c(0.2, 0.8))

roll_sum <- rollify(
  ~sum(.x),
  window = 120
)

spy_returns %>% 
  mutate(regime_1_prob =  mix_mod$posterior[,1],
         regime_2_prob =  mix_mod$posterior[,2],
         bear_regime =  case_when(round(mix_mod$posterior[,2]) == 1 ~ 1,
                                  TRUE ~ 0),
         rolling_bear_obs = roll_sum(bear_regime),
         recession_prob = sigmoid(scale(rolling_bear_obs))) %>% 
  na.omit() %>%
  ggplot(aes(x = date)) +
  geom_line(aes(y = recession_prob), color = "cornflowerblue") +
  labs(y = "recession prob") +
  scale_y_continuous(labels = scales::percent)
```


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

